---
title: "HIV-Ninis"
subtitle: "Exploratory data analysis"
author: Diego Villa
date: "`r Sys.time()`"
output: html_notebook
---

# Packages

```{r}
library(fs)
library(readr)
library(dplyr)
library(ggplot2)
library(scales)
library(sf)
library(srvyr)
library(scales)
```

# Data loading

```{r}
#TODO Set column types
dataset <- read_csv('../../data/processed/dataset.csv')
```

# Data validation

Check that there is no duplicates in the data.

```{r}
dataset %>% 
  group_by(year) %>% 
  summarise(n = n(), n_caseid = length(unique(caseid)))
```

# Exploratory data analysis

## HIV screening by age groups and wealth index through the years

```{r}
#TODO Change age to age_women

dataset %>% 
  filter(age >= 15) %>%  
  mutate(
    wealth_index = factor(
      wealth_index, 
      c('Poorest', 'Poor', 'Middle', 'Rich', 'Richest')
    ),
    natural_region = factor(
      natural_region,
      c('Lima Metropolitan Area', 'Rest of Coast', 'Highland', 'Jungle')
    ),
    highest_edu_level = factor(
      highest_edu_level,
      c('No education', 'Primary', 'Secondary', 'Higher')
    )
  ) -> dataset_analysis
```

```{r}
dataset_analysis %>% 
  mutate(weighting_factor_scaled = weighting_factor / 1000000) %>% 
  as_survey_design(
    ids = conglomerate, strata = stratum, weight = weighting_factor_scaled, 
    nest = TRUE
  ) -> data_analysis_srvy
```

```{r}
data_analysis_srvy %>% 
  filter(age <= 29) %>% 
  # group_by(year) %>% 
  mutate(
    age_group = cut(
      age,
      quantile(age), include.lowest = TRUE, right = FALSE
    ),
    nini = if_else(
      currently_attending_school == "No" & respondent_currently_working == "No",
      "Yes", "No"
    )
  ) %>% 
  mutate(
    wealth = factor(
      case_when(
        wealth_index == "Poorest" ~ "Poor",
        wealth_index == "Poor" ~ "Poor",
        wealth_index == "Middle" ~ "Middle",
        wealth_index == "Rich" ~ "Rich",
        wealth_index == "Richest" ~ "Rich"
      ),
      levels = c("Poor", "Middle", "Rich")
    )
  ) %>% 
  mutate(
    nini_wealth = factor(
      case_when(
        nini == "Yes" & wealth == "Poor" ~ "Nini: Poor",
        nini == "Yes" & wealth == "Middle" ~ "Nini: Middle",
        nini == "Yes" & wealth == "Rich" ~ "Nini: Rich",
        nini == "No" & wealth == "Poor" ~ "No-nini: Poor",
        nini == "No" & wealth == "Middle" ~ "No-nini: Middle",
        nini == "No" & wealth == "Rich" ~ "No-nini: Rich"
      ),
      levels = c(
        "Nini: Poor", "Nini: Middle", "Nini: Rich", "No-nini: Poor", 
        "No-nini: Middle", "No-nini: Rich"
      )
    )
  ) ->
  data_ninis_srvy
```

```{r}
data_ninis_srvy %>% 
  filter(
    across(c(nini_wealth, tested_aids_hiv_12_months), ~ !is.na(.x)),
    nini == "Yes"
  ) %>% 
  group_by(age_group, year, nini_wealth) %>% 
  summarise(
    prop_screen = survey_mean(
      tested_aids_hiv_12_months == "Yes", vartype = "ci", proportion = TRUE
    ),
    .groups = "drop"
  ) ->
  data_plot_1
```

```{r fig.width=6, fig.height=12}
data_plot_1 %>% 
  ggplot(aes(year, prop_screen, color = nini_wealth)) +
  geom_errorbar(
    aes(ymin = prop_screen_low, ymax = prop_screen_upp), width = 0.25, 
    linetype = "dashed"
  ) +
  geom_line() +
  geom_point(size = 4) +
  scale_y_continuous(labels = percent) +
  theme_bw() +
  theme(legend.position = "top") +
  facet_wrap(~ age_group, ncol = 1)
```


Check number of cases by age group.

```{r}
dataset_analysis %>% 
  group_by(age_group) %>% 
  summarise(cases = n())
```

```{r}
# dataset_analysis %>% 
#   filter(tested_aids_hiv_12_months %in% c("Yes", "No")) -> 
#   dataset_analysis_screen
```

```{r}
# dim(dataset_analysis_screen)
```

```{r}
# dataset_analysis_screen %>% 
#   group_by(year) %>% 
#   summarise(n = n(), n_caseid = length(unique(caseid)))
```

Raw number and proportion of women in every age group that answered or not if they got tested for HIV in the last 12 months at the time of the survey in each year.

```{r}
dataset_analysis %>% 
  group_by(year, age_group, tested_aids_hiv_12_months) %>% 
  summarise(cases = n()) %>% 
  mutate(prop_cases = cases / sum(cases))
```
```{r}
# dataset_analysis %>% 
#   mutate(weighting_factor_scaled = weighting_factor / 1000000) %>% 
#   as_survey_design(
#     ids = conglomerate, strata = stratum, weight = weighting_factor_scaled, 
#     nest = TRUE
#   ) -> data_analysis_srvyr

dataset_analysis %>% 
  # filter(across(tested_aids_hiv_12_months, ~ !is.na(.x))) %>% 
  mutate(weighting_factor_scaled = weighting_factor / 1000000) %>% 
  as_survey_design(
    ids = conglomerate, strata = stratum, weight = weighting_factor_scaled, 
    nest = TRUE
  ) -> data_analysis_srvyr
```

```{r}
data_analysis_srvyr %>% 
  filter(across(tested_aids_hiv_12_months, ~ !is.na(.x))) %>% 
  group_by(year, age_group, tested_aids_hiv_12_months) %>% 
  summarise(
    prop_cases = survey_mean(vartype = "ci", proportion = TRUE)
  ) %>% 
  ungroup() -> 
  screening_year_age_groups
```

```{r fig.height=10}
screening_year_age_groups %>% 
  ggplot(
    aes(
      x = age_group, y = prop_cases, fill = tested_aids_hiv_12_months, 
      ymax = prop_cases_upp, ymin = prop_cases_low  
    )
  ) +
  geom_col(position = "dodge") +
  geom_errorbar(position = "dodge") + 
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~ year, ncol = 2) +
  theme_minimal() +
  theme(legend.position = "top")
```

```{r}
dataset_analysis %>% 
  group_by(tested_aids_hiv_12_months, currently_attending_school) %>% 
  summarise(cases = n()) %>% 
  mutate(prop_cases = cases / sum(cases))
```


```{r}
data_analysis_srvyr %>% 
  mutate(
    nini = if_else(
      currently_attending_school == 'No' & respondent_currently_working == 'No',
      "Yes", "No"
    )
  ) %>% 
  group_by(year, age_group, tested_aids_hiv_12_months) %>% 
  summarise(
    prop_cases = survey_mean(vartype = "ci", proportion = TRUE)
  ) %>% 
  ungroup() -> 
  screening_year_age_groups
```


```{r}
dataset_analysis %>% 
  filter(age_group != "[30,49]") %>% 
  group_by(ever_been_tested_aids) %>% 
  summarise(cases = n()) %>% 
  mutate(prctg = 100 * cases / sum(cases))
```


```{r}
dataset_analysis %>% 
  group_by(diagnosted_std_last_12_months) %>% 
  summarise(cases = n()) %>% 
  mutate(prctg = 100 * cases / sum(cases))
```

```{r}
dataset_analysis %>% 
  group_by(diagnosted_std_last_12_months_cs) %>% 
  summarise(cases = n()) %>% 
  mutate(prctg = 100 * cases / sum(cases))
```

```{r}
dataset_analysis %>% 
  group_by(respondent_currently_working, type_earnings_work) %>% 
  summarise(cases = n()) %>% 
  mutate(prctg = 100 * cases / sum(cases))
```


```{r}
dataset_analysis %>% 
  group_by(respondent_currently_working, work_last_12_months) %>% 
  summarise(cases = n()) %>% 
  mutate(prctg = 100 * cases / sum(cases))
```


```{r}
dataset_eda <- 
  dataset %>% 
  mutate(
    nini = 
      ifelse(
        (age >= 15) & (age <= 24) & (currently_attending_school == 'No') &
          (respondent_currently_working == 'No'),
        'Yes', 'No'
      ),
    wealth_index = factor(
      wealth_index, 
      c('Poorest', 'Poor', 'Middle', 'Rich', 'Richest')
    ),
    natural_region = factor(
      natural_region,
      c('Lima Metropolitan Area', 'Rest of Coast', 'Highland', 'Jungle')
    ),
    highest_edu_level = factor(
      highest_edu_level,
      c('No education', 'Primary', 'Secondary', 'Higher')
    ),
    year = factor(
      year,
      as.character(2015:2020)
    )
  ) 
```

```{r}
table(dataset_eda$nini, useNA = "ifany")
100 * table(dataset_eda$nini, useNA = "ifany") / nrow(dataset_eda)
```

```{r}
dataset_eda <- filter(dataset_eda, !is.na(nini))
```

```{r}
dataset_eda %>% 
  group_by(wealth_index, nini) %>% 
  summarize(cases = n()) %>% 
  mutate(prop = cases / sum(cases)) %>% 
  ungroup() %>% 
  ggplot(aes(wealth_index, prop, fill = nini)) +
  geom_bar(stat = "identity", position = "stack") + 
  scale_y_continuous(labels = percent)
```

```{r}
dataset_eda %>% 
  group_by(natural_region, nini) %>% 
  summarize(cases = n()) %>% 
  mutate(prop = cases / sum(cases)) %>% 
  ungroup() %>% 
  ggplot(aes(natural_region, prop, fill = nini)) +
  geom_bar(stat = "identity", position = "stack") + 
  scale_y_continuous(labels = percent)
```

```{r}
dataset_eda %>% 
  group_by(region, nini) %>% 
  summarize(cases = n()) %>% 
  mutate(prop = cases / sum(cases)) %>% 
  ungroup() %>% 
  ggplot(aes(region, prop, fill = nini)) +
  geom_bar(stat = "identity", position = "stack") + 
  scale_y_continuous(labels = percent) +
  coord_flip()
```

```{r}
dataset_eda %>% 
  group_by(highest_edu_level, nini) %>% 
  summarize(cases = n()) %>% 
  mutate(prop = cases / sum(cases)) %>% 
  ungroup() %>% 
  ggplot(aes(highest_edu_level, prop, fill = nini)) +
  geom_bar(stat = "identity", position = "stack") + 
  scale_y_continuous(labels = percent)
```

```{r}
table(dataset_eda$ever_been_tested_aids, useNA = 'ifany')
100 * table(dataset_eda$ever_been_tested_aids, useNA = 'ifany') / nrow(dataset_eda)
```

```{r}
dataset_eda %>% 
  group_by(nini, ever_been_tested_aids) %>% 
  summarize(cases = n()) %>% 
  mutate(prop = cases / sum(cases)) %>% 
  ungroup() %>% 
  ggplot(aes(nini, prop, fill = ever_been_tested_aids)) +
  geom_bar(stat = "identity", position = "stack") + 
  scale_y_continuous(labels = percent)
```

```{r}
dataset_eda %>% 
  filter(!is.na(diagnosed_std_last_12_months)) %>% 
  group_by(nini, diagnosed_std_last_12_months) %>% 
  summarize(cases = n()) %>% 
  mutate(prop = cases / sum(cases)) %>% 
  ungroup() %>% 
  ggplot(aes(nini, prop, fill = diagnosed_std_last_12_months)) +
  geom_bar(stat = "identity", position = "stack") + 
  scale_y_continuous(labels = percent)
```

```{r}
dataset_eda %>% 
  group_by(region, ever_been_tested_aids) %>% 
  summarize(cases = n()) %>% 
  mutate(prop = cases / sum(cases)) %>% 
  ungroup() %>% 
  ggplot(aes(region, prop, fill = ever_been_tested_aids)) +
  geom_bar(stat = "identity", position = "stack") + 
  scale_y_continuous(labels = percent) +
  coord_flip()
```

```{r fig.height=3, fig.width=10}
dataset_eda %>% 
  group_by(wealth_index, nini, ever_been_tested_aids) %>% 
  summarize(cases = n()) %>% 
  mutate(prop = cases / sum(cases)) %>% 
  ungroup() %>% 
  ggplot(aes(nini, prop, fill = ever_been_tested_aids)) +
  facet_wrap(vars(wealth_index), ncol = 6) +
  geom_bar(stat = "identity", position = "stack") + 
  scale_y_continuous(labels = percent)
```

```{r}
dataset_eda %>% 
  group_by(year, nini) %>% 
  summarize(cases = n()) %>% 
  mutate(prop = cases / sum(cases)) %>% 
  ungroup() %>% 
  ggplot(aes(year, prop, fill = nini)) +
  geom_bar(stat = "identity", position = "stack") + 
  scale_y_continuous(labels = percent)
```

```{r}
dataset_eda %>% 
  group_by(year, ever_been_tested_aids) %>% 
  summarize(cases = n()) %>% 
  mutate(prop = cases / sum(cases)) %>% 
  ungroup() %>% 
  ggplot(aes(year, prop, fill = ever_been_tested_aids)) +
  geom_bar(stat = "identity", position = "stack") + 
  scale_y_continuous(labels = percent)
```















